---
title: 'Week 3: Polling and Elections'
author: "Janet Hernandez"
date: "2022-09-25"
slug: []
categories: []
tags: Prediction
summary: This weeks blog post will focus on incorporating polling data to add into my existing predictive model that already includes economy variables to test to see how our predictions improve.
---
```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(scales)
library(plotly)
library(rjson)
library(jtools)
library(htmlwidgets)
library(reactable)
library(ggplot2)
library(ggrepel)
```

```{r, include=FALSE}
CPI_monthly <- read.csv("~/Desktop/Gov1347 Data/Section data 2/CPI_monthly.csv")
economy <- read.csv("~/Desktop/Gov1347 Data/Section data 2/GDP_quarterly.csv")
unemployment_state_monthly <- read_csv("~/Desktop/Gov1347 Data/Section data 2/unemployment_state_monthly.csv")
popvote_df <- read_csv("~/Desktop/Gov1347 Data/Section data 2/house_popvote_seats.csv")
unemployment_national_quarterly_final <- read_csv("~/Desktop/Gov1347 Data/Section data 2/unemployment_national_quarterly_final.csv")
RDI_quarterly <- read_csv("~/Desktop/Gov1347 Data/Section data 2/RDI_quarterly.csv")
RDI_monthly <- read_csv("~/Desktop/Gov1347 Data/Section data 2/RDI_monthly.csv")
voteseatshare_2020 <- read.csv("~/Desktop/Syllabi for Fall 22/election analytics/Gov1347 Data/house nationwide vote and seat share by party 1948-2020.csv")
house_party_vote_share_by_district_1948_2020 <- read_csv("~/Desktop/Syllabi for Fall 22/election analytics/Gov1347 Data/house party vote share by district 1948-2020.csv")
economy_q2 <- economy %>% 
  filter(quarter_yr == 2) 

RDI_q2 <- RDI_quarterly %>%
  filter(quarter_yr == 2)

unemployment_q2 <- unemployment_national_quarterly_final %>%
  filter(quarter_yr == 2)

economy_model_data <- popvote_df %>% 
  inner_join(economy_q2, by = "year")

economy_model_data <-economy_model_data %>% 
  inner_join(RDI_q2, by = "year")

economy_model_data <-economy_model_data %>% 
  inner_join(unemployment_q2, by = "year")

library(grid)

theme_janet <- function(base_size = 13, base_family = "") {
  theme_grey(base_size = base_size, base_family = base_family) %+replace%
    theme(
      
      # Base elements which are not used directly but inherited by others
      line =              element_line(colour = '#DADADA', size = 0.75, 
                                       linetype = 1, lineend = "butt"),
      rect =              element_rect(fill = "#F0F0F0", colour = "#F0F0F0", 
                                       size = 0.5, linetype = 1),
      text =              element_text(family = base_family, face = "plain",
                                       colour = "#656565", size = base_size,
                                       hjust = 0.5, vjust = 0.5, angle = 0, 
                                       lineheight = 0.9),
      
      # Modified inheritance structure of text element
      plot.title =        element_text(size = rel(1.5), family = '' , 
                                       face = 'bold', hjust = -0.05, 
                                       vjust = 1.5, colour = '#3B3B3B'),
      axis.title.x =      element_text(),
      axis.title.y =      element_text(),
      axis.text =         element_text(),
      
      # Modified inheritance structure of line element
      axis.ticks =        element_line(),
      panel.grid.major =  element_line(),
      panel.grid.minor =  element_blank(),
      
      # Modified inheritance structure of rect element
      plot.background =   element_rect(),
      panel.background =  element_rect(),
      legend.key =        element_rect(colour = '#DADADA'),
      
      # Modifiying legend.position
      legend.position = 'none',
      
      complete = TRUE
    )
}
```
# Introduction # 

This week, we focused on seeing how polling might affect and predict election outcomes. [Gelman and King](https://www-jstor-org.ezp-prod1.hul.harvard.edu/stable/194212?sid=primo#metadata_info_tab_contents) spoke about how polling is really only indicative of people's preferences close to the election. Over the course of a campaign cycle, Gelman and King argue, responses to pollsters during the campaign are not generally informed or even, in a sense we describe, "rational." 

For this reason, when building and assessing a model this week, I will try to weigh the polls heavier when they are closer to the election day. 

# My Predictive Model So Far #

To clarify from my previous posts where I didn't clearly have a section of my model and its predictions formally, I will do that now. I have decided to look at the Democratic seat number as the dependent variable for my model. My previous model combined economy data which I chose to be GDP, disposable income, and unemployment. With this model in my previous post, the R-squared is 0.22. This is relatively good, considering that 20% of the variance can be explained by these trends. However, the residual standard error is 29.18 on 27 degrees of freedom. Below is my a graph of my model that includes my observed vs predicted values. 


```{r, echo=FALSE}

PredictiveModel <- lm(D_seats ~  GDP_growth_qt + DSPIC_change_qt + UNRATE, data = economy_model_data)

summ(PredictiveModel)


National_prediction <- data.frame()

data_mod <- data.frame(Predicted = predict(PredictiveModel, new_data = National_prediction),  # Create data for ggplot2 
                       Observed = economy_model_data$D_seats)


```

```{r plot for prediction, echo= FALSE}
ggplot(data_mod,                                  
       aes(x = Predicted,
           y = Observed))  +
  geom_point() + 
  geom_abline(intercept = 0,
              slope = 1,
              color = "red",
              size = 2) +
 geom_label_repel(label= economy_model_data$year, nudge_y = 6) + theme_janet() + ggtitle("Predictive Model 1:\nPredicted v. Observed Democrat Seats") 
```
**While the model is not perfect, I am impressed that the 2020 election was almost perfectly predicted as well as 1972 and 1968. However, the most recent election years in the past two decades my model has primarily been over predicting Dem seats. This is clearly seen in years like 2016, 2014, and 2012.**

```{r, actual prediction for 2022 election}

predict(PredictiveModel, data.frame(GDP_growth_qt = 496.200, DSPIC_change_qt = -929.06
, UNRATE = 3.4))

```

My current model's prediction *without* polling information is that for the upcoming election, the Dems will win approximately **255** seats in Congress. One thing to note is that I used last quarter's inflation rate because our current DSPIC change was so extreme it resulted in an implausible number, so instead I used Q1 DSPIC change which is more in-line with typical rates as my figure for prediction in the model.

# Introduction to Polling #

This week, I will look at how polling data might affect our model. I will be doing extension #1, which is discussing differences between the 538 and Economist models and then building my own model again to predict the election with the added variable of polling. I will weigh this as (365 - days until the election).

# Economist vs. Five Thirty Eight Prediction Models #


```{r, include=FALSE}
# Read in data
pollster_ratings <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/pollster-ratings/2020/pollster-ratings.csv")
poll_df <- read.csv("~/Downloads/GenericPolls1942_2020 (1).csv")

```

```{r clean and add weights, include=FALSE}
poll_df <- poll_df %>%
  # Filter out election year and discard polls conducted after election
  filter(year %% 2 == 0 & year >= 1948) %>%
  filter(days_until_election > 0, days_until_election < 365) %>%
  select(pollster, year, emonth, eday, dem, rep, days_until_election) %>%
  rename(dem_poll = dem, rep_poll = rep) %>%
  # add weights
  mutate(recency = 365 - days_until_election)

# add a column that includes the actual seat count in poll? 


#nvm found this gonna use this lol 

results <- read.csv("~/Downloads/Gov1374Fall2022CongressionalSeatChanges - Sheet1.csv")

#Combine our results data and then the poll data 
new <- left_join(poll_df, results, by =c("year"= "ElectionDate"))


```

```{r weighted data, include = FALSE}
polling_final <- poll_df %>%
  # weighted average
  group_by(year) %>%
  summarize(dem_poll_weighted = weighted.mean(dem_poll, recency),
            rep_poll_weighted = weighted.mean(rep_poll, recency)) %>%
  rename(election_year = year)

# combine new weighted polling data with my previous dataframe for predictions

polling_new <- left_join(economy_model_data, polling_final, by = c("year"= "election_year"))

```

# New Model Results #
```{r weighted regression with polling, echo=FALSE}
newmodel <- lm(D_seats ~  GDP_growth_qt + DSPIC_change_qt + UNRATE + dem_poll_weighted, data = polling_new)
summ(newmodel)

prediction2 <- data.frame()

pred2 <- data.frame(Predicted = predict(newmodel, new_data = prediction2),  # Create data for ggplot2 
                       Observed = polling_new$D_seats)


ggplot(pred2,                                  
       aes(x = Predicted,
           y = Observed))  +
  geom_point() + 
  geom_abline(intercept = 0,
              slope = 1,
              color = "red",
              size = 1) +
 geom_label_repel(label= polling_new$year, nudge_y = 6) + theme_janet() + ggtitle("Predictive Model 2:\nPredicted v. Observed Democrat Seats") 

```
My new model has a much better R^2 when I consider the weighted polling results. It also seems to be a much closer fit to the observed values with few exceptions like 2008. I think this is indicative that weighted polling results close to the election do help us improve our predictions about elections. 



## References
Abramowitz, A. (2018). Will Democrats Catch a Wave? The Generic Ballot Model and the 2018 US House Elections. *PS: Political Science & Politics, 51*(S1), 4-6. doi:10.1017/S1049096518001567

Bafumi, J., Erikson, R., & Wlezien, C. (2018). Forecasting the 2018 Midterm Election using National Polls and District Information. *PS: Political Science & Politics, 51*(S1), 7-11. doi:10.1017/S1049096518001579


Gelman, A., & King, G. (1993). Why Are American Presidential Election Campaign Polls So Variable When Votes Are So Predictable? *British Journal of Political Science, 23*(4), 409–451. https://doi.org/10.1017/S0007123400006682

